---
title:   "A doublesex model for reduced dimorphism - Normalizing the raw metabolome data"
output:   html_notebook
by:       R. Coig
updated:  8/16/24
---
Notebook Summary: Normalizing the raw metabolome data
```{r,load required libraries}
library(dplyr)
library(reshape2)
```
Reading metabolomics data from csv files.
```{r,read in raw data}
rm(list=ls())
data<-read.csv("../_data/raw/raw_metabolome.csv")
key<-read.csv("../_data/sample_key.csv")
```
91 metabolites are represented with no missing data.
```{r}
data<-merge(key,data,by=c("Sample"))
data<-data[!duplicated(data),]#removing the duplicate data for N-METHYLGLUTAMATE
data$Sample<-factor(data$Sample)#converting Sample to factor
data$Metabolite<-factor(data$Metabolite)#converting Metabolite to factor
mz<-unique(data$Metabolite)#91 metabolites
sum(is.na(data$Normalized.Area))#no missing data
```
Create a matrix of all sample readings w/ Metabolites as row names, and individual matrices for each of the tissue types (head, thorax, abdomen)
```{r}
# Filter, select, reshape the data
matrix_samples_all <- data %>%
  filter(Type != "standard") %>%
  select(Name, Metabolite, Normalized.Area) %>%
  dcast(Metabolite ~ Name, value.var = "Normalized.Area")

# Set row names and extract metabolites and sample names
row.names(matrix_samples_all) <- matrix_samples_all$Metabolite
mz <- matrix_samples_all$Metabolite
matrix_samples_all$Metabolite <- NULL
samples_all <- colnames(matrix_samples_all)#51 total samples

#XXGA4 was reported as an outlier, outlier test in another notebook
matrix_samples <- matrix_samples_all%>%select(-XXGA4)
samples <- colnames(matrix_samples)#50 samples after removing outlier
```
Splitting dfs based on the 4th letter of each column name which is Tissue Type for tissue-specific analyses
```{r}
matrix_Head <- matrix_samples %>% select(starts_with(substring(colnames(matrix_samples), 1, 3) %>% paste0("H")))
matrix_Thorax <- matrix_samples %>% select(starts_with(substring(colnames(matrix_samples), 1, 3) %>% paste0("T")))
matrix_Abdomen <- matrix_samples %>% select(starts_with(substring(colnames(matrix_samples), 1, 3) %>% paste0("A")))
matrix_Abdomen_all <- matrix_samples_all %>% select(starts_with(substring(colnames(matrix_samples_all), 1, 3) %>% paste0("A")))

# Extract column names for samples
samples_H <- colnames(matrix_Head)
samples_T <- colnames(matrix_Thorax)
samples_A <- colnames(matrix_Abdomen)
samples_A_all <- colnames(matrix_Abdomen_all)
```
Log transform, scale and center each dataset by column for pca and analyses
```{r}
# Defining a function for log scaling and centering
process_matrix <- function(matrix) {
  as.data.frame(scale(log(matrix), center = TRUE, scale = TRUE))
}

# Applying the function to each matrix
mz_log_forpca <- process_matrix(matrix_samples)
mz_log_forpca_all <- process_matrix(matrix_samples_all)
mz_log_forpca_head <- process_matrix(matrix_Head)
mz_log_forpca_thorax <- process_matrix(matrix_Thorax)
mz_log_forpca_abdomen <- process_matrix(matrix_Abdomen)
mz_log_forpca_abdomen_all <- process_matrix(matrix_Abdomen_all)
```
Saving out centered and scaled R data files, for outlier test and further analysis
```{r,save out r files}
save(file='../_data/scaled.data', mz_log_forpca, mz_log_forpca_head, mz_log_forpca_thorax, mz_log_forpca_abdomen, mz, samples)
save(file='../_data/scaled_with_outlier.data', mz_log_forpca_all, mz_log_forpca_head, mz_log_forpca_thorax, mz_log_forpca_abdomen_all, mz, samples)
save(file='../_data/keys.data', key, mz)
rm(list=ls())
```