---
title:   "A doublesex model for reduced dimorphism - ANOVA and Tukey HSD"
output:   html_notebook
by:       R. Coig
updated:  8/28/24
---
Notebook Summary: ANOVA and within-tissue testing for significant differences in metabolite levels between wildtype or dsx null sexes
```{r,load required libraries}
library(dplyr)
```
Loading in data
```{r,loading data}
rm(list = ls())
load('../_data/scaled.data')
load('../_data/abdomen_with_outlier_removed.data')
load('../_data/keys.data')
```
Formatting data frames for significance testing
```{r}
data_Head<-as.data.frame(t(mz_log_forpca_head))
data_Thorax<-as.data.frame(t(mz_log_forpca_thorax))
data_Abdomen<-as.data.frame(t(mz_log_forpca_abdomen))
data_Abdomen_noXXwA4<-as.data.frame(t(mz_log_forpca_abdomen_noXXwA4))
row.names(key)<-key$Name
```
adding group labels to dfs
```{r}
source("src/mergegroups_function.R")
data_Head<-mergegroups(data_Head)
data_Thorax<-mergegroups(data_Thorax)
data_Abdomen<-mergegroups(data_Abdomen)
data_Abdomen_noXXwA4<-mergegroups(data_Abdomen_noXXwA4)
```
Tukey's HSD testing within each tissue df
```{r}
source("src/Tukeytesting_loop.R")
data_list<-list(data_Head, data_Thorax, data_Abdomen, data_Abdomen_noXXwA4)
data_Tukey_tested <- lapply(data_list, Tukeytesting) 
names(data_Tukey_tested) <- c("Tukey_Head_test","Tukey_Thorax_test","Tukey_Abdomen_test","Tukey_Abdomen_test_noXXwA4")
list2env(data_Tukey_tested,globalenv())
```
Combining results for all tissues into one df, adding tissue type
```{r}
Tukey_all_tested<-rbind(Tukey_Head_test,Tukey_Thorax_test,Tukey_Abdomen_test)
Tukey_all_tested$Ttype<-c(rep("Head",91),rep("Thorax",91),rep("Abdomen",91))
Tukey_all_tested<-Tukey_all_tested %>%select(1, Ttype, everything())

Tukey_all_tested_noXXwA4<-rbind(Tukey_Head_test,Tukey_Thorax_test,Tukey_Abdomen_test_noXXwA4)
Tukey_all_tested_noXXwA4$Ttype<-c(rep("Head",91),rep("Thorax",91),rep("Abdomen",91))
Tukey_all_tested_noXXwA4<-Tukey_all_tested_noXXwA4 %>%select(1, Ttype, everything())
```
Correcting for multiple comparisons
```{r}
source("src/multiple_testing_correction.R")
Tukey_all_tested<-multiple_testing(Tukey_all_tested)
Tukey_all_tested_noXXwA4<-multiple_testing(Tukey_all_tested_noXXwA4)
```
Save out results:
```{r,save out results in csv format}
write.csv(Tukey_all_tested,"output/02a_Tukey_test_statistics.csv",row.names = F)
write.csv(Tukey_all_tested_noXXwA4,"output/02a_Tukey_test_statistics_Abdomen_no_outlier.csv",row.names = F)
```
Saving out .data files
```{r,save out r files}
save(file='../_results/Tukey_test_results.data', Tukey_all_tested, Tukey_all_tested_noXXwA4)
rm(list = ls())
```