---
title:   "A doublesex model for reduced dimorphism - SD tests"
output:   html_notebook
by:       R. Coig
updated:  8/28/24
---
Notebook Summary: Statistical tests for significant reduction of sex differences
```{r,load required libraries}
library(dplyr)
library(reshape2)
```
Loading in data
```{r,loading data}
rm(list = ls())
load('../_results/Tukey_test_results.data')
```
Calculating total unique number of sig metabolites
```{r}
unique(Tukey_all_tested%>%filter(sig_w=="*")%>%dplyr::select(Metabolite))->sig_mzs
Tukey_all_tested%>%filter(Metabolite%in%sig_mzs$Metabolite)%>%select(Metabolite,Ttype,ES_w,ES_d,FDR_w,FDR_d)->sigatleastone
```
Significant metabolites by tissue for wildtype and dsx null flies
```{r}
unique(Tukey_all_tested%>%filter(FDR_w<.05&Ttype=="Head")%>%dplyr::select(Metabolite))->sigHeadmz
unique(Tukey_all_tested%>%filter(FDR_w<.05&Ttype=="Thorax")%>%dplyr::select(Metabolite))->sigThoraxmz
unique(Tukey_all_tested%>%filter(FDR_w<.05&Ttype=="Abdomen")%>%dplyr::select(Metabolite))->sigAbdomenmz
unique(Tukey_all_tested_noXXwA4%>%filter(FDR_w<.05&Ttype=="Abdomen")%>%dplyr::select(Metabolite))->sigAbdomenmz_noXXwA4
sig_w_head<-unique(as.character(sigHeadmz$Metabolite))
sig_w_thorax<-unique(as.character(sigThoraxmz$Metabolite))
sig_w_abdomen<-unique(as.character(sigAbdomenmz$Metabolite))
sig_w_abdomen_noXXwA4<-unique(as.character(sigAbdomenmz_noXXwA4$Metabolite))
Tukey_all_tested%>%filter(FDR_d<.05&Ttype=="Head")%>%dplyr::select(Metabolite)%>%pull()%>%as.character()->sig_dsx_head
Tukey_all_tested%>%filter(FDR_d<.05&Ttype=="Thorax")%>%dplyr::select(Metabolite)%>%pull()%>%as.character()->sig_dsx_thorax
Tukey_all_tested%>%filter(FDR_d<.05&Ttype=="Abdomen")%>%dplyr::select(Metabolite)%>%pull()%>%as.character()->sig_dsx_abdomen
```
Data for density plots
```{r}
dist_melt<-melt(Tukey_all_tested%>%select("Metabolite","Ttype","absES_w","absES_d"),id.vars=c("Metabolite","Ttype"))
dist_melt_sig<-melt(Tukey_all_tested%>%filter(Metabolite%in%sigatleastone$Metabolite)%>%dplyr::select("Metabolite","Ttype","absES_w","absES_d"),id.vars=c("Metabolite","Ttype"))

#mean and median of sex effect sizes across all metabolites
ES_allmz_summary <- dist_melt%>%group_by(Ttype,variable)%>%
  summarise(mean=round(mean(value),4),
            median=median(value))
#mean and median of sex effect sizes across SD metabolites
ES_SDmz_summary <- dist_melt_sig%>%group_by(Ttype,variable)%>%
  summarise(mean=round(mean(value),4),
            median=median(abs(value)))
```
Formatting Table 1
```{r}
Table1<-merge(ES_SDmz_summary[,1:3],ES_allmz_summary[,1:3],by=c("variable","Ttype"))%>%
  rename(SDGroup=variable,SD_meanES=mean.x,meanES=mean.y)
Table1$N<-c(1,1,1,24,22,33)
Table1$percent<-round((Table1$N/91),2)*100
```
Formatting df to plot diffs for 3G and VSC flies by tissue type
```{r}
Tukey_w<-Tukey_all_tested%>%dplyr::select("Metabolite","Ttype","ES_w","FDR_w","absES_w")
colnames(Tukey_w)<-c("Metabolite","Ttype","ES","p","absES")
Tukey_w$SDgroup<-"w"
Tukey_d<-Tukey_all_tested%>%dplyr::select("Metabolite","Ttype","ES_d","FDR_d","absES_d")
colnames(Tukey_d)<-c("Metabolite","Ttype","ES","p","absES")
Tukey_d$SDgroup<-"d"
Tukmelt<-rbind(Tukey_w,Tukey_d)
Tukmelt$log10<-log10(Tukmelt$absES)
```
Filtering by those Higher in XX, versus those higher in XY
```{r}
Tukey_all_tested%>%filter(sig_w=="*"&ES_w<0,Ttype=="Head")%>%dplyr::select(Metabolite)->sigHeadmzXX
Tukey_all_tested%>%filter(sig_w=="*"&ES_w>0,Ttype=="Head")%>%dplyr::select(Metabolite)->sigHeadmzXY
Tukey_all_tested%>%filter(sig_w=="*"&ES_w<0,Ttype=="Thorax")%>%dplyr::select(Metabolite)->sigThoraxmzXX
Tukey_all_tested%>%filter(sig_w=="*"&ES_w>0,Ttype=="Thorax")%>%dplyr::select(Metabolite)->sigThoraxmzXY
Tukey_all_tested%>%filter(sig_w=="*"&ES_w<0,Ttype=="Abdomen")%>%dplyr::select(Metabolite)->sigAbdomenmzXX
Tukey_all_tested%>%filter(sig_w=="*"&ES_w>0,Ttype=="Abdomen")%>%dplyr::select(Metabolite)->sigAbdomenmzXY
```
Formatting df to plot diffs for 3G and VSC flies by tissue type
```{r}
Tukmelt<-Tukmelt[!duplicated(Tukmelt),]
Tukmelt%>%filter(Ttype=="Head"&Metabolite%in%sigHeadmzXX$Metabolite)->TukmeltsigHeadXX
Tukmelt%>%filter(Ttype=="Thorax"&Metabolite%in%sigThoraxmzXX$Metabolite)->TukmeltsigThoraxXX
Tukmelt%>%filter(Ttype=="Abdomen"&Metabolite%in%sigAbdomenmzXX$Metabolite)->TukmeltsigAbdomenXX
TukmeltplotXX<-rbind(TukmeltsigHeadXX,TukmeltsigThoraxXX,TukmeltsigAbdomenXX)

Tukmelt%>%filter(Ttype=="Head"&Metabolite%in%sigHeadmzXY$Metabolite)->TukmeltsigHeadXY
Tukmelt%>%filter(Ttype=="Thorax"&Metabolite%in%sigThoraxmzXY$Metabolite)->TukmeltsigThoraxXY
Tukmelt%>%filter(Ttype=="Abdomen"&Metabolite%in%sigAbdomenmzXY$Metabolite)->TukmeltsigAbdomenXY
TukmeltplotXY<-rbind(TukmeltsigHeadXY,TukmeltsigThoraxXY,TukmeltsigAbdomenXY)
```
Fig 2B and 2C, Metabolites with sig sex difference in 3G, but no sex difference in VSC flies
```{r}
dat_text_XX <- data.frame(
  label = c("n = 10", "n = 15", "n = 11"),
  Ttype   = factor(c("Head","Thorax","Abdomen"))
)
dat_text_XY <- data.frame(
  label = c("n = 12", "n = 18", "n = 13"),
  Ttype   = factor(c("Head","Thorax","Abdomen"))
)
Table2<-merge(dat_text_XX,dat_text_XY,by="Ttype")%>%
  rename(HigherXX=label.x,HigherXY=label.y)
```
Statistical test comparing global sex differences in the metabolome for each tissue, saving out p-values for plot descriptions. Across ALL metabolites (Figure 2B)
```{r}
Tukmelt_Head<-Tukmelt%>%filter(Ttype=="Head")
fit_H<-lm(data=Tukmelt_Head,log10~SDgroup)
Headresults<-anova(fit_H)
p_head_allmz<-round(Headresults$`Pr(>F)`[1],3)

Tukmelt_Thorax<-Tukmelt%>%filter(Ttype=="Thorax")
fit_T<-lm(data=Tukmelt_Thorax,log10~SDgroup)
Thoraxresults<-anova(fit_T)
p_thorax_allmz<-round(Thoraxresults$`Pr(>F)`[1],3)

Tukmelt_Abdomen<-Tukmelt%>%filter(Ttype=="Abdomen")
fit_A<-lm(data=Tukmelt_Abdomen,log10~SDgroup)
Abdomenresults<-anova(fit_A)
p_abdomen_allmz<-round(Abdomenresults$`Pr(>F)`[1],4)
```
Statistical test comparing global sex differences in the metabolome for each tissue, saving out p-values for plot descriptions. Just for SD metabolites (Figure 2A)
```{r}
Tukmelt_Head<-Tukmelt%>%filter(Ttype=="Head")%>%filter(Metabolite%in%sigatleastone$Metabolite)
fit_H<-lm(data=Tukmelt_Head,log10~SDgroup)
Headresults<-anova(fit_H)
p_head_SDmz<-round(Headresults$`Pr(>F)`[1],3)

Tukmelt_Thorax<-Tukmelt%>%filter(Ttype=="Thorax")%>%filter(Metabolite%in%sigatleastone$Metabolite)
fit_T<-lm(data=Tukmelt_Thorax,log10~SDgroup)
Thoraxresults<-anova(fit_T)
p_thorax_SDmz<-round(Thoraxresults$`Pr(>F)`[1],4)

Tukmelt_Abdomen<-Tukmelt%>%filter(Ttype=="Abdomen")%>%filter(Metabolite%in%sigatleastone$Metabolite)
fit_A<-lm(data=Tukmelt_Abdomen,log10~SDgroup)
Abdomenresults<-anova(fit_A)
p_abdomen_SDmz<-round(Abdomenresults$`Pr(>F)`[1],6)
```
Save out Table 1 and 2:
```{r,save out results in csv format}
write.csv(Table1,"output/02b_Table1.csv",row.names = F)
write.csv(Table2,"output/02b_Table2.csv",row.names = F)
```
Saving out .data files
```{r,save out r files}
save(file='../_results/Figure2.data', dist_melt, dist_melt_sig, ES_allmz_summary, ES_SDmz_summary, p_abdomen_allmz,p_abdomen_SDmz,p_thorax_allmz,p_thorax_SDmz,p_head_allmz,p_head_SDmz)
save(file='../_results/Figure3.data', sig_w_abdomen, sig_w_head, sig_w_thorax)
rm(list = ls())
```